---
title: "Capstone Intro to NFL data"
author: "JMeerse"
date: "`r Sys.Date()`"
output: html_document
---

```{r}
```

# Getting started with NFL data and R

Start by loading the nflfastR package and library, and tidyverse. You only have to do this once. (This is already set up for you!)

```{r setup}
knitr::opts_chunk$set(echo = TRUE)
# install.packages("nflfastR",repos = "http://cran.us.r-project.org")
# install.packages("tidyverse",repos = "http://cran.us.r-project.org")
```

## Load libraries

R makes "packages" of commands that are useful for different kinds of data analysis. In order to use those commands, you load the libraries of commands *each time* you run your project.

```{r libraries}
library(nflfastR)
library(tidyverse)
```

## Load data

NflfastR allows us to download all the plays from various seasons. Let's get all the plays from the 2021-22 season, into a dataframe called pbp21.

```{r data}
pbp21 <- load_pbp(2021) 
```

### What did we just do?

We used the command `load_pbp()` to specify we wanted the play-by-play data from the 2021-22 season.\
We also assigned that play-by-play data to a dataframe (similar to a spreadsheet) called pbp21.\

In the upper right of your RStudio window you should see the name of the dataframe. How many plays were run during the season?

Now let's look at the data.

```{r looking at our data}
head(pbp21)

#names(pbp21)

```

The first command, `head()`, showed the first 6 rows of the data, and only 6 of the 372 variables recorded for each play.\
The second command, `names()`, will show the names of all 372 variables.  How do we know what each variable represents?

Go to the link: <https://nflreadr.nflverse.com/articles/dictionary_pbp.html>

## Initial Questions - using `filter()` to just look at runs 

The command below takes the pbp21 dataframe and applies 2 filters, and creates a new dataframe.\

1)  What is the new dataframe called?\
2)  What games are represented in the dataframe?\
3)  What type of plays are in the dataframe?\
4) Run the code -- how many running plays are recorded?  

```{r initial questions}
run21 <- pbp21 %>% filter(season_type == "REG" & play_type == "run") 
```

Let's look at yards gained on run plays using statistics and graphs.

```{r summary stats and graphs}
summary(run21$yards_gained) #summary stats

run21 %>% ggplot(aes(x = yards_gained)) + 
  geom_boxplot() #boxplot of all runs - use + facet_wrap(~posteam) to see each team


run21 %>% ggplot(aes(x = yards_gained)) + 
  geom_histogram() 
```

Let's compare the mean and median run for each team.

We'll use `group_by()` to look at results by team, and `summarise()` to compute these statistics.

```{r average by team}
run_by_team <- run21 %>% group_by(posteam) %>% 
  summarise(mean_run = mean(yards_gained, na.rm = TRUE),
            median_run = median(yards_gained, na.rm = TRUE),
            n = n()) %>% 
  arrange(desc(n)) %>% 
  ungroup()
```

Let's graph the teams by mean yards gained.

```{r a graph of mean yards by team}

run_by_team %>% 
  ggplot(aes(x = reorder(posteam, -mean_run), y = mean_run)
         ) + 
  geom_col()
```

Not bad, but let's clean things up a bit.

First, let's switch the axes to see team names easier.
```{r a nicer graph}
run_by_team %>% 
  ggplot(aes(x = reorder(posteam, mean_run), y = mean_run)
         ) + 
  coord_flip() +
  geom_col()
```

Now, let's get team colors! 

First we have to add on to the dataframe.
```{r adding colors}
run_by_team <- left_join(run_by_team, teams_colors_logos, by = c("posteam" = "team_abbr"))
```

Now let's add color to the graph:
```{r color by team}
run_by_team %>% 
  ggplot(aes(x = reorder(posteam, mean_run), y = mean_run)
         ) + 
  coord_flip() +
  geom_col(color = run_by_team$team_color2,
           fill = run_by_team$team_color)
```

Now, go back up to line 67, copy all the way to line 131, paste it below and rework the code to investigate passing yards.

Good luck!

